---
title: "Homework 4"
author: "Christopher Brunswick"
date: "11/1/2021"
output: word_document
---
1. Use the $ operator to access the population size data and store it
as the object pop. Then use the sort function to redefine pop so that
it is sorted. Finally, use the [ operator to report the smallest
population size.
```{r}
library(dslabs)
data(murders)
murders$population
pop <- murders$population
srtpop <- sort(pop)
srtpop[1]
```

2. Now instead of the smallest population size, find the index 
of the entry with the smallest population size. Hint: use order
instead of sort

```{r}
index_small <- order(pop)
index_small[1]
```
3. We can actually perform the same operation as in the previous
exercise using the function which.min. Write one line of code that
does this.


```{r}
which.min(pop)
```
4. Now we know how small the smallest state is and we know which 
row represents it. Which state is it? Define a variable states
to be the state names from the murders data frame. Report the name
of the state with the smallest population.

```{r}
states <- murders$state
min <- which.min(pop)
states[min]
```
5. You can create a data frame using the data.frame function.
Use the rank function to determine the population rank of each
state from smallest population size to biggest. Save these ranks
in an object called ranks, then create a data frame with the state
name and its rank. Call the data frame my_df.

```{r}
ranks <- rank(pop)
my_df <- data.frame(state_name=states, pop_ranks=ranks)
my_df
```


6. Repeat the previous exercise, but this time order my_df so that the states are ordered from least populous to most populous. Hint:
create an object ind that stores the indexes needed to order the
population values. Then use the bracket operator [ to re-order
each column in the data frame.

```{r}
ind <- order(pop)
my_df <- data.frame(state_name=states[ind], pop_rank=sort(ranks))
my_df
```


7. The na_example vector represents a series of counts. You
can quickly examine the object using:

data("na_example")
str(na_example)
mean(na_example)

The is.na function returns a logical vector that tells us which
entries are NA. Assign this logical vector to an object called ind
and determine how many NA's does na_example have.
```{r}
is.na(na_example)
ind <- is.na(na_example)
na_example[ind]
length(na_example[ind])
```


8. Now compute the average again, but only for the entries that
are not NA. Hint: remember the ! operator.
```{r}
na.omit(na_example)
no_na <- na.omit(na_example)
mean(no_na)
```
