---
title: "Homework 7"
author: "Christopher Brunswick"
date: "11/10/2021"
output: word_document
---
1. We made a plot of total murders versus population and noted a strong relationship. Not surprisingly, states with larger populations had more murders. 
```{r}
library(dslabs)
data(murders)
population_in_millions <- murders$population/10^6
total_gun_murders <- murders$total
plot(population_in_millions, total_gun_murders)
```

keep in mind that many states have populations below 5 million and
are bunched up. We may gain further insights from making this plot 
in the log scale. Transform the variables using the log10
transformation and then plot them.

```{r}
plot(log10(population_in_millions), log10(total_gun_murders))
```

2. Create a histogram of the state populations

```{r}
state_pop <- with(murders, murders$population)
hist(state_pop)
```

3. Generate boxplots of the state populations by region

```{r}
boxplot(population~region, data = murders, xlab="region", ylab="population")
```
