---
title: "Homework 9"
author: "Christopher Brunswick"
date: "11/23/2021"
output: word_document
---
1. Load the dplyr package and the murders dataset.
```{r}
library(dplyr)
library(dslabs)
data(murders)
```

# You can add columns using the dplyr function mutate. 
# This function is aware of the column names and inside 
# the function you can call them unquoted:


```{r}
murders <- mutate(murders, population_in_millions = population / 10^6)
murders
```

# We can write population rather than murders$population. 
#The function mutate knows we are grabbing columns from murders.

```{r}
murders <- mutate(murders, rate = total/population * 10^6 , population_in_millions = population / 10^6)
murders
```

# 2. If rank(x) gives you the ranks of x from lowest to highest, 
# rank(-x) gives you the ranks from highest to lowest. 
# Use the function mutate to add a column rank containing the rank, 
# from highest to lowest murder rate. Make sure you redefine murders 
# so we can keep using this variable.

```{r}
murders <- mutate(murders, rate = total / population * 100000, rank = rank(-rate) , population_in_millions = population / 10^6)
murders
```


# 3. With dplyr, we can use select to show only certain columns. 
# For example, with this code we would only show the states and 
# population sizes:

```{r}
select(murders, state, population) %>% head()
```

# Use select to show the state names and abbreviations in murders. 
# Do not redefine murders, just show the results.

```{r}
select(murders, state, abb) %>% head()
```


# 4. The dplyr function filter is used to choose specific rows of 
# the data frame to keep. Unlike select which is for columns, 
# filter is for rows. For example, you can show just the New York 
# row like this:

```{r}
filter(murders, state == "New York")
```

# Use filter to show the top 5 states with the highest murder rates. 
# After we add murder rate and rank, do not change the murders 
# dataset, just show the result. Remember that you can filter based 
# on the rank column.

```{r}
filter(murders, rank <= 5)
```


# 5. We can remove rows using the != operator. For example, to 
# remove Florida, we would do this:

```{r}
no_florida <- filter(murders, state != "Florida")
no_florida
```


# Create a new data frame called no_south that removes states from 
# the South region. How many states are in this category? You can 
# use the function nrow for this.

```{r}
my_table <- select(murders, state, region)
no_south <- filter(my_table, region != "South")
nrow(no_south)
```

# 6. We can also use %in% to filter with dplyr. 
# You can therefore see the data from New York and Texas like this:

```{r}
filter(murders, state %in% c("New York", "Texas"))
```

# Create a new data frame called murders_nw with only the states 
# from the Northeast and the West. How many states are in this 
# category?

```{r}
murders_nw <- filter(my_table, region %in% c("Northeast", "West"))
murders_nw
nrow(murders_nw)
```


# 7. Suppose you want to live in the Northeast or West and want 
# the murder rate to be less than 1. We want to see the data for 
# the states satisfying these options. Note that you can use 
# logical operators with filter. Here is an example in which we 
# filter to keep only small states in the Northeast region.

```{r}
filter(murders, population < 5000000 & region == "Northeast")
```


# Make sure murders has been defined with rate and rank and 
# still has all states. Create a table called my_states that 
# contains rows for states satisfying both the conditions: it is in 
# the Northeast or West and the murder rate is less than 1. Use 
# select to show only the state name, the rate, and the rank.

```{r}
my_states <- filter(murders, rate < 1 & region %in% c("Northeast", "West"))
my_states
select(my_states, state, rate, rank)
```

# 8. The pipe %>% can be used to perform operations sequentially 
# without having to define intermediate objects. Start by redefining 
# murder to include rate and rank.

```{r}
murders <- mutate(murders, rate =  total / population * 100000, 
                  rank = rank(-rate))
```

# In the solution to the previous exercise, we did the following:

```{r}
my_states <- filter(murders, region %in% c("Northeast", "West") & 
                        rate < 1)
select(my_states, state, rate, rank)
```


# The pipe %>% permits us to perform both operations sequentially 
# without having to define an intermediate variable my_states. 
# We therefore could have mutated and selected in the same line 
# like this:

```{r}
mutate(murders, rate =  total / population * 100000, 
       rank = rank(-rate)) %>%
  select(state, rate, rank)
```

# Notice that select no longer has a data frame as the first 
# argument. The first argument is assumed to be the result of 
# the operation conducted right before the %>%.

# Repeat the previous exercise, but now instead of creating a new 
# object, show the result and only include the state, rate, and rank
# columns. Use a pipe %>% to do this in just one line.

```{r}
mutate(murders, region %in% c("Northeast", "West") & rate < 1) %>% select(state, rate, rank)
```

# 9. Reset murders to the original table by using data(murders). 
# Use a pipe to create a new data frame called my_states that 
# considers only states in the Northeast or West which have a 
# murder rate lower than 1, and contains only the state, rate and 
# rank columns. The pipe should also have four components separated 
# by three %>%. The code should look something like this:

```{r}
my_states <- murders %>% mutate(rate= total / 
              population*100000, rank=rank(-rate)) %>% 
  filter(region %in% c("Northeast", "West") & rate < 1) %>%
  select(state, rate, rank)
my_states
```
