---
title: "Homework 5"
author: "Christopher Brunswick"
date: "11/2/2021"
output: word_document
---
1. Previously we created this data frame:

temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
city_temps <- data.frame(name=city, temperature=temp)

Remake the data frame using the code above, but add a line that converts
the temperature from fahrenheit to celsius. The conversion is 
C = 5/9 * (F-32)
```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
newtemp <- 5/9*(temp-32)
city_temps <- data.frame(name=city, temperature=newtemp)
city_temps
```

2. What is the following sum 1 + 1/2^2 + 1/3^3 + ...+ 1/100^2?
Hint: thanks to Euler, we know it should be close to pi^2/6

```{r}
n <- 1:100
q <- 1/n^2
sum(q)
```

3. Compute the per 100,000 murder rate for each state and store it
in the object murder_rate. Then compute the average murder rate for
the US using the function mean. What is the average?

```{r}
library(dslabs)
data(murders)
str(murders)
murders$total/100000
murder_rate <- murders$total/murders$population*100000
mean(murder_rate)
```


