---
title: "Homework 8"
author: "Christopher Brunswick"
date: "11/16/2021"
output: word_document
---
# 1. What will this conditional expression return?
```{r}
library(dslabs)
data(murders)
x <- c(1,2,-3,4)
if(all(x>0)){
  print("All Postives")
} else{
  print("Not all positives")
}
```

# 2. Which of the following expressions is always FALSE when at 
# least one entry of a logical vector x is TRUE?

```{r}
any(!x)
```

# 3. The function nchar tells you how many characters 
# long a character vector is. 
# Write a line of code that assigns to the object new_names the 
# state abbreviation when the state name is longer than 8 characters.


```{r}
state_name <- murders$state
length_of_name <- nchar(state_name) > 8
murders$abb[length_of_name]
new_names <- murders$abb[length_of_name]
```

# 4. Create a function sum_n that for any given value, say n, computes 
# the sum of the integers from 1 to n (inclusive). 
# Use the function to determine the sum of integers from 1 to 5,000.

```{r}
sum_n <- function(n){ 
  x <- 1:n  
  print(sum(x))
  }
sum_n(5000)
```


# 5. Create a function altman_plot that takes 
# two arguments, x and y, and plots the difference against the sum.

```{r}
altman_plot <- function(x,y){
  print(plot(x-y,x+y))
}
altman_plot(10,100)
```

# 6. After running the code below, what is the value of x?

```{r}
x <- 3
my_func <- function(y){
  x <- 5
  y+5
}
my_func(x)
```

# 7. Write a function compute_s_n that for any given n computes 
# the sum Sn=12+22+32+…n2. Report the value of the sum when n=10.

```{r}
compute_s_n <- function(n){
  x <- 1:n
  q <- x^2
  sum(q)
}
compute_s_n(10)
```

# 8. Define an empty numerical vector s_n of size 25 
# using s_n <- vector("numeric", 25) and store in the results 
# of S1,S2,…S25 using a for-loop.

```{r}
m <- 25
s_n <- vector( length = m)
for (n in 1:m){
  s_n[n] <- compute_s_n(n)
}
s_n
```

# 9. Repeat exercise 8, but this time use sapply.

```{r}
n <- 1:25
sapply(n , compute_s_n)
```

# 10. Plot Sn versus n. Use points defined by n=1,…,25.

```{r}
plot(n,s_n)
```


# 11. Confirm that the formula for this sum is Sn=n(n+1)(2n+1)/6.

```{r}
t <- 1:25
q <- (t*(t+1)*(2*t+1))/6
q
```
